Flowchart of Current Algorithm

### 2009-01-20 docs/flowchart/flowchart.tex
If use_tfromp=T, we pass temperature through react_state. If use_tfromp=F, we update temperature in react_state using the eos.
working on "Updates" Section

@@

\item {\bf React State}$[\rho^{\inp},(\rho h)^{\inp},X_k^{\inp},T^{\inp}, (\rho\Hext)^{\inp}]\rightarrow [\rho^{\outp}, (\rho h)^{\outp}, X_k^{\outp}, T^{\outp}, (\rho \omegadot_k)^{\outp}, (\rho\Hnuc)^{\outp}]$ \\
is the process by which we evolve the species and enthalpy from 
$X_k^{\inp} \rightarrow X_k^{\outp}$ 
and $(\rho h)^{\inp} \rightarrow (\rho h)^{\outp}$ by solving the following 
system of equations over a time interval of  $\Delta t/2$,
\begin{eqnarray}
\frac{\partial X_k}{\partial t} &=& \omegadot_k\enskip,\\
\frac{\partial (\rho h)}{\partial t} &=& \rho \Hnuc + \rho \Hext\enskip .
\end{eqnarray}
  In particular, to evolve the species, we solve the system:
\begin{eqnarray}
\frac{dX_k}{dt} &=& \omegadot_k(\rho,X_k,T)\enskip, \\
\frac{dT}{dt} &=&\frac{1}{c_p} \left ( -\sum_k \xi_k  \omegadot_k  + \Hnuc \right )
\enskip. \label{eq:reacttemp}
\end{eqnarray}
using the stiff ordinary differential equation integration methods
provided by the {\tt VODE} package (cite VODE).  The absolute error
tolerances are set to $10^{-12}$ for the species, and a relative
tolerance of $10^{-5}$ is used for the temperature.  The integration
yields the new values of the mass fractions, $X_k^{\outp}$.  Equation
(\ref{eq:reacttemp}) is derived from the full temperature evolution equation
\begin{equation}
\label{eq:tempenthalpy}
\frac{DT}{Dt} = \frac{1}{\rho c_p} \left[ \left(1 - \rho h_p\right) \frac{D p}{D t}
 - \sum_k \rho \xi_k {\omegadot}_k 
 + \rho \Hnuc + \rho \Hext \right] \enskip , 
\end{equation}
(paper III; eq.~[8])
by assuming that the pressure is constant during the burn state.  In
evolving these equations, we need to evaluate $c_p$ and $\xi_k$.  In
theory, this means evaluating the equation of state for each
right-hand side evaluation that {\tt VODE} requires.  In practice, we
freeze $c_p$ and $\xi_k$ at the start of the integration time step and
compute them using $\rho^{\inp}, X_k^{\inp},$ and $T^{\inp}$ as inputs
to the equation of state.  We note that while temperature is evolved
when solving these equations, we do not keep the final temperature,
nor do we use it to compute the final change in enthalpy.  Therefore,
$T^{\outp} = T^{\inp}$.  Also, note that the density remains unchanged
during the {\bf React State} step, i.e., $\rho^{\outp} = \rho^{\inp}$.
\MarginPar{Currently we do update the temperature.}

@@

We now describe the full time advancement algorithm, making frequent
use of the shorthand developed above.  Here, we assume that the
problem is already properly initialized.  We describe the details of
the initialization in \S~\ref{Sec:Initialization}.

There are several changes to the algorithm since it was last detailed
in paper III.  The major changes are:
\begin{itemize}
\item In paper III we discretized the enthalpy evolution equation in
full, non-perturbation form.  Since then we have discovered that 
discretizing the enthalpy evolution in perturbational form leads to better
numerical properties.  This is more like paper II.

\item We added the volume-discrepancy term to the velocity constraint
equation.
\end{itemize}
%
The advance of the state through a single timestep appears as:

@@

  \begin{enumerate}
  \item Perform steps 1--11 as described above, using 
    $S^{\myhalf,\star} = (S^0 + S^{1,\nu})/2$ in step 1 as described.
    The only other difference in the time advancement is that in step 11
    we define ${\bf V} = (\Ubt^{1,\star} - \Ubt^0)$ and solve
    \begin{equation}  L_\beta^\rho \phi =
      D \left ( \beta_0^{\myhalf} {\bf V} \right) - \beta_0^{\myhalf} \left[ (S^{1}-\overline{S}^{1}) - (S^{0}-\overline{S}^{0}) \right] \enskip . 
    \end{equation}
    (The motivation for this form of the projection in the initial pressure iterations
    is discussed in (cite almgren:bell:crutchfield).)
      We discard the new velocity resulting from this, but keep the new  
      value for $\pi^{\myhalf} = \pi^{-\myhalf} + (1 / \dt) \; \phi.$  
      These steps also yield new scalar data at time $\dt,$ which
      we discard,  and new values for $\etarho^{\myhalf}$ (step 8), $\psi^{\myhalf}$ (step 8), 
      $S^{1,\nu+1}$ (step 10), and $\pi^{\myhalf}$ (step 11), which we keep.
    \item Set $\pi^{-\myhalf} = \pi^{\myhalf}$, $\etarho^{-\myhalf} = \etarho^{\myhalf}$,
      and $\psi^{-\myhalf} = \psi^{\myhalf}$. 
    \end{enumerate}

@@

In summary:
\begin{equation}
\left.\begin{array}{ccccccccc}
& \text{React State} & & \text{Advect State} & & \text{Diffuse Enthalpy} & & \text{React State} & \\
\rho^n & & & \longrightarrow & & & & & \rho^{n+1} \\
(\rho X)^n & \longrightarrow & (\rho X)^{(1)} & \longrightarrow & & & (\rho X)^{(2)} & \longrightarrow & (\rho X)^{n+1} \\
(\rho h)^n & \longrightarrow & (\rho h)^{(1)} & \longrightarrow & (\rho h)^{(2a)} & \longrightarrow & (\rho h)^{(2)} & \longrightarrow & (\rho h)^{n+1}\\
T^n & \longrightarrow & T^{(1)} & & & \longrightarrow & T^{(2)} & \longrightarrow & T^{n+1} \\
\end{array}\right.\nonumber
\end{equation}
\begin{equation}
\left.\begin{array}{ccccccccc}
& \text{React Base} & & \text{Advect Base} & & \text{Correct Base} & & \text{React Base} & \\
\rho_0^n & & & \longrightarrow & \rho_0^{(2a)} & \longrightarrow & & & \rho_0^{n+1} \\
(\rho h)_0^n & \longrightarrow & (\rho h)_0^{(1)} & \longrightarrow & & & (\rho h)_0^{(2)} & \longrightarrow & (\rho h)_0^{n+1} \\
p_0^n & & & \longrightarrow & & & & & p_0^{n+1} \\
\end{array}\right.
\end{equation}
Note that this chart indicates, for example, that $\rho^n = \rho^{(1)}$, $\rho^{(2a)} = \rho^{(2)} = \rho^{n+1}$, etc.

>>>>

\section{Algorithm Updates}

There are several changes to the algorithm since it was last detailed
in paper III.  The major changes are:

\begin{itemize}
\item In paper III we discretized the enthalpy evolution equation in
full, non-perturbation form.  Since then we have discovered that 
discretizing the enthalpy evolution in perturbational form leads to better
numerical properties.  This is more like paper II.

\item We split {\bf Advect Base} into separate {\bf Advect Base Density},
{\bf Advect Base Pressure}, and {\bf Advect Base Enthalpy}.  This is because we use
the results from {\bf Advect Base Density} to compute a time-centered $\etarho$ which 
is used as an input to {\bf Advect Base Pressure}.  Also, we use the results from
{\bf Advect Base Pressure} to compute a time-centered $\psi$ which is used as an input
to {\bf Advect Base Enthalpy}.

\item We added the volume-discrepancy term to the velocity constraint
equation.
\end{itemize}

@@

\item {\bf React State}$[\rho^{\inp},(\rho h)^{\inp},X_k^{\inp},T^{\inp}, (\rho\Hext)^{\inp}]\rightarrow [\rho^{\outp}, (\rho h)^{\outp}, X_k^{\outp}, T^{\outp}, (\rho \omegadot_k)^{\outp}, (\rho\Hnuc)^{\outp}]$ \\
is the process by which we evolve the species and enthalpy from 
$X_k^{\inp} \rightarrow X_k^{\outp}$ 
and $(\rho h)^{\inp} \rightarrow (\rho h)^{\outp}$ by solving the following 
system of equations over a time interval of  $\Delta t/2$,
\begin{eqnarray}
\frac{\partial X_k}{\partial t} &=& \omegadot_k\enskip,\\
\frac{\partial (\rho h)}{\partial t} &=& \rho \Hnuc + \rho \Hext\enskip .
\end{eqnarray}
  In particular, to evolve the species, we solve the system:
\begin{eqnarray}
\frac{dX_k}{dt} &=& \omegadot_k(\rho,X_k,T)\enskip, \\
\frac{dT}{dt} &=&\frac{1}{c_p} \left ( -\sum_k \xi_k  \omegadot_k  + \Hnuc \right )
\enskip. \label{eq:reacttemp}
\end{eqnarray}
using the stiff ordinary differential equation integration methods
provided by the {\tt VODE} package (cite VODE).  The absolute error
tolerances are set to $10^{-12}$ for the species, and a relative
tolerance of $10^{-5}$ is used for the temperature.  The integration
yields the new values of the mass fractions, $X_k^{\outp}$.  Equation
(\ref{eq:reacttemp}) is derived from the full temperature evolution equation
\begin{equation}
\label{eq:tempenthalpy}
\frac{DT}{Dt} = \frac{1}{\rho c_p} \left[ \left(1 - \rho h_p\right) \frac{D p}{D t}
 - \sum_k \rho \xi_k {\omegadot}_k 
 + \rho \Hnuc + \rho \Hext \right] \enskip , 
\end{equation}
(paper III; eq.~[8])
by assuming that the pressure is constant during the burn state.  In
evolving these equations, we need to evaluate $c_p$ and $\xi_k$.  In
theory, this means evaluating the equation of state for each
right-hand side evaluation that {\tt VODE} requires.  In practice, we
freeze $c_p$ and $\xi_k$ at the start of the integration time step and
compute them using $\rho^{\inp}, X_k^{\inp},$ and $T^{\inp}$ as inputs
to the equation of state.  We note that while temperature is evolved
when solving these equations, we do not keep the final temperature,
nor do we use it to compute the final change in enthalpy.  Therefore,
$T^{\outp} = T^{\inp}$.  Also, note that the density remains unchanged
during the {\bf React State} step, i.e., $\rho^{\outp} = \rho^{\inp}$.
\MarginPar{Currently we do update the temperature if {\tt use\_tfromp}=F .}

@@

We now describe the full time advancement algorithm, making frequent
use of the shorthand developed above.  Here, we assume that the
problem is already properly initialized.  We describe the details of
the initialization in \S~\ref{Sec:Initialization}.
The advance of the state through a single timestep appears as:

@@

  \begin{enumerate}
  \item Perform steps 1--11 as described above, using 
    $S^{\myhalf,\star} = (S^0 + S^{1,\nu})/2$ in step 1 as described.
    The only other difference in the time advancement is that in step 11
    we define ${\bf V} = (\Ubt^{1,\star} - \Ubt^0)$ and solve
    \begin{equation}  L_\beta^\rho \phi =
      D \left ( \beta_0^{\myhalf} {\bf V} \right) - \beta_0^{\myhalf} \left[ \left(S^{1}-\overline{S}^{1}\right) - \left(S^{0}-\overline{S}^{0}\right) \right] \enskip . 
    \end{equation}
    (The motivation for this form of the projection in the initial pressure iterations
    is discussed in (cite almgren:bell:crutchfield).)
      We discard the new velocity resulting from this, but keep the new  
      value for $\pi^{\myhalf} = \pi^{-\myhalf} + (1 / \dt) \; \phi.$  
      These steps also yield new scalar data at time $\dt,$ which
      we discard,  and new values for $\etarho^{\myhalf}$ (step 8), $\psi^{\myhalf}$ (step 8), 
      $S^{1,\nu+1}$ (step 10), and $\pi^{\myhalf}$ (step 11), which we keep.
    \item Set $\pi^{-\myhalf} = \pi^{\myhalf}$, $\etarho^{-\myhalf} = \etarho^{\myhalf}$,
      and $\psi^{-\myhalf} = \psi^{\myhalf}$. 
    \end{enumerate}

@@

In summary:
\begin{equation}
\left.\begin{array}{ccccccccc}
& \text{React State} & & \text{Advect State} & & \text{Diffuse Enthalpy} & & \text{React State} & \\
\rho^n & & & \longrightarrow & & & & & \rho^{n+1} \\
(\rho X_k)^n & \longrightarrow & (\rho X_k)^{(1)} & \longrightarrow & & & (\rho X_k)^{(2)} & \longrightarrow & (\rho X_k)^{n+1} \\
(\rho h)^n & \longrightarrow & (\rho h)^{(1)} & \longrightarrow & (\rho h)^{(2a)} & \longrightarrow & (\rho h)^{(2)} & \longrightarrow & (\rho h)^{n+1}\\
T^n & \longrightarrow & T^{(1)} & & & \longrightarrow & T^{(2)} & \longrightarrow & T^{n+1} \\
\end{array}\right.\nonumber
\end{equation}
\begin{equation}
\left.\begin{array}{ccccccccc}
& \text{React Base} & & \text{Advect Base} & & \text{Correct Base} & & \text{React Base} & \\
\rho_0^n & & & \longrightarrow & \rho_0^{(2a)} & \longrightarrow & & & \rho_0^{n+1} \\
(\rho h)_0^n & \longrightarrow & (\rho h)_0^{(1)} & \longrightarrow & & & (\rho h)_0^{(2)} & \longrightarrow & (\rho h)_0^{n+1} \\
p_0^n & & & \longrightarrow & & & & & p_0^{n+1} \\
\end{array}\right.
\end{equation}
Note that this chart indicates, for example, that $\rho^n = \rho^{(1)}$, $\rho^{(2a)} = \rho^{(2)} = \rho^{n+1}$, etc.
